The Impact of Anthropogenic Forcing on ENSO Amplitude

### 2021-12-21 presentations/presentation/presentation.tex
some changes lol

\begin{frame}{Review of Literature}
  \begin{itemize}
  \item ENSO's properties observed vary across different decades. \citep{lubbecke2014assessing}.
  \item ENSO responds to external forcing.
    \begin{itemize}
    \item Correlation between ENSO strength and sunspot activity \citep{emile2007nino}.
    \item Weakened ENSO during the Ice Age due to reduced CO$_2$ levels \citep{zhu2017reduced}.
    \end{itemize}
  \item Models show possible increasing ENSO activity in the future \citep{zheng2017response} and \citep{maher2018enso}.
  \item Factors other than CO\textsubscript{2} can affect ENSO.
    \begin{itemize}
    \item Ozone emissions reduce ENSO activity \citep{nowack2017role}.
    \item Aerosol emissions modify ENSO geographical center \citep{stevenson2017forced}.
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Gap}
  \begin{itemize}
  \item Little research using a large ensemble to examine the effect of individual factors on ENSO.
  \item Considerable disagreement between studies on whether ENSO will strengthen or weaken due to global warming
  \end{itemize}
\end{frame}

\begin{frame}{Questions}
  \begin{description}
  \item[What?] Do the CESM1 and CESM2 predict increased or decreased ENSO intensity in the future?
  \item[Why?] Is the predicted increase (or decrease) due to human activities?
  \item[How?] What processes are causing greenhouse gasses and aerosols to affect ENSO?
  \end{description}
\end{frame}

\begin{frame}{Methods Overview}
  \begin{itemize}
  \item Precollected predictions of sea surface temperature from climate models.
  \item Calculate ENSO intensity in model output.
  \item Use single forcing ensembles to estimate contributions each forcing item.
  \item Plot correlation between ENSO intensity and ocean temperature to examine relationship between heat transfer, forcing, and ENSO.
  \item Use wavelet analysis to analyze changes to ENSO at different frequencies.
  \end{itemize}
\end{frame}

\begin{frame}{Role of Mentor and Student}
  \begin{columns}[t]
    \column{.5\textwidth}
    Student:
    \begin{itemize}
    \item Analyze raw data on computer
    \item Produce graphics for analysis and publication
    \item Write documentation
    \item Identify key features of results
    \end{itemize}
    \column{.5\textwidth}
    Mentor:
    \begin{itemize}
    \item Review student writing
    \item Interpret results in the context of climatology
    \item Conduct parallel analysis
    \item Provide raw data from facility
    \end{itemize}
  \end{columns}

\end{frame}

\begin{frame}{Model Setup}
  \begin{itemize}
  \item CESM1 \citep{kay2015community} and CESM2 \citep{danabasoglu2020community}
  \item Observed forcing levels from 1850-2005
  \item Predicted forcing levels from 2005-2100
  \item Ensembles have 40 and 50 simulations respectively
  \item Control simulation with pre-1850 forcing levels
  \item Single forcing ensembles that represent influence of single factor
  \end{itemize}
\end{frame}

@@

\begin{frame}{Discussion}
  \begin{itemize}
  \item Rising greenhouse gas levels increase Pacific Ocean stratification, strengthening ENSO cycle.
  \item Aerosol influence nonlinear because aerosol levels are not purely increasing.
  \item Stronger ENSO will lead to greater climate variability and extreme weather.
  \item
  \end{itemize}
\end{frame}

\begin{frame}{Conclusions and Discussion}
  \begin{itemize}
  \item There is likely to be an increase in ENSO strength over the next 100 years. Agrees with \citet{cai2018increased}.
  \item Variance increase is likely caused by the combined influence of greenhouse gasses and aerosols.
  \item Global warming increases ENSO intensity by warming upper layers of the Pacific faster than central layers.
  \end{itemize}
\end{frame}

\begin{frame}{Application, Limitation, and Next Steps}
  \begin{description}
  \item[\alert{Application}] Improve prediction ability to help people prepare for increased likelihood of extreme weather.
  \item[\alert{Limitation}] Niño 3.4 index may not be fully accurate for various models \citep{cai2018increased}. Also, CESM may contain biases and is not completely accurate.
  \item[\alert{Next steps}] Examine other variables to further analyze mediator process, continue wavelet analysis methods to focus on individual frequency bands.
  \end{description}
\end{frame}

@@

>>>>

\begin{frame}{Review of Literature}
  \begin{itemize}
  \item ENSO's properties observed vary across different decades. \citep{lubbecke2014assessing}.
  \item ENSO responds to external forcing.
  \item Weakened ENSO during the Ice Age due to reduced CO$_2$ levels \citep{zhu2017reduced}.
  \item Models show possible increasing ENSO activity in the future \citep{zheng2017response} and \citep{maher2018enso}.
  \end{itemize}
\end{frame}

\begin{frame}{Gap and Questions}
  \begin{itemize}
  \item Little research using a large ensemble to examine the effect of individual factors on ENSO.
  \item Considerable disagreement between studies on whether ENSO will strengthen or weaken due to global warming
  \end{itemize}
  \begin{description}
  \item[What?] Do the CESM1 and CESM2 predict increased or decreased ENSO intensity in the future?
  \item[Why?] Is the predicted increase (or decrease) due to human activities?
  \item[How?] What processes are causing greenhouse gasses and aerosols to affect ENSO?
  \end{description}
\end{frame}

\begin{frame}{Model Setup (Data)}
  \begin{itemize}
  \item CESM1 \citep{kay2015community} and CESM2 \citep{danabasoglu2020community}
  \item Predicts climate over 21st century with global warming.
  \item 40-50 simulations per ensemble.
  \item Control simulation with pre-1850 forcing levels.
  \item Single forcing ensembles that represent influence of single factor.
  \end{itemize}
\end{frame}

@@

\begin{frame}{Discussion}
  \begin{itemize}
  \item Rising greenhouse gas levels increase Pacific Ocean stratification, strengthening ENSO cycle.
  \item Aerosol influence is nonlinear because aerosol levels are not purely increasing.
  \item Stronger ENSO may lead to greater temperature variability and extreme weather.
  \item CESM1 and CESM2 conflict in their prediction of the changes to ENSO's frequency.
  \end{itemize}
\end{frame}

\begin{frame}{Limitations and Applications}
  Limitations:
  \begin{itemize}
  \item Niño 3.4 index shown to be inaccurate for some models \citep{cai2018increased}.
  \item CESM may contain biases.
  \end{itemize}
  Application: to improve our ability to predict ENSO and help people prepare for increased likelihood of extreme weather.
\end{frame}

@@

\begin{frame}{Role of Mentor and Student}
  \begin{columns}[t]
    \column{.4\textwidth}
    Student:
    \begin{itemize}
    \item Analyze raw data on computer
    \item Produce graphics for analysis and publication
    \item Write documentation
    \item Identify key features of results
    \end{itemize}
    \column{.4\textwidth}
    Mentor:
    \begin{itemize}
    \item Review student writing
    \item Interpret results in the context of climatology
    \item Conduct parallel analysis
    \item Provide raw data from facility
    \end{itemize}
  \end{columns}
\end{frame}
